### 2018-06-22 pubs/itp-2018/sections/proof-reconstruction/resolve.tex
[ pubs ] preparation of the paper (#60)
* [ paper ] removed many examples and extra content.

* [ paper ] all proofs have been removed.

* Complete first revision to section about Metis

* Review section 4.2

* Review introduction Section 4

* Improve many parts in the text.

Metis section improves in the tptp description.
The example tstp derivation was added with line
numbers.
The explanation of the workflow has changed to
be more explict and also improved for readability.
Remove the numbers in the diagram.
Appendix updated removing the tstp of the small example.

* [ Sec4 ] example of the s

Logic equivalence between propositions is a major issue to justify
prover reasoning steps. Since we left out semantics to treat only the
syntax aspects of propositional logic, our approach shows logic
equivalence by converting propositions to their conjunctive normal
form, and reordering those and the inner disjunctions to match them.

@@

A \emph{literal} is an propositional variable (positive literal) or a
negation of an propositional variable (negative literal).

@@

\begin{notation}
We use \Lit as synonym of \Prop type to refer literals.
\end{notation}

@@

The \emph{conjunctive normal form} of a formula also called clausal
normal form is the logical equivalent version expressed as a
conjunction of clauses where a \emph{clause} is the disjunction of
zero or more literals.

\begin{myexamplenum}\hspace{10cm}
\label{ex:resolve-tstp}
\begin{verbatim}
  cnf(r₄, ¬ r ∨ p ∨ q, ...
  cnf(r₅, p ∨ q ∨ r, ...
  cnf(r₆, p ∨ q, inf(resolve, r, [r₅, r₄])).
\end{verbatim}

In the excerpt above, we apply resolution to the first two formulas,
$¬ r ∨ p ∨ q$ and $p ∨ q ∨ r$.
The last line tells us the literal used
for resolution is $r$. Syntactically speaking, we can not derive
neither the conclusion $p ∨ q$ in \verb!r₆! nor apply the resolution
theorem with \verb!r₄! and \verb!r₅! since the formulas do not fit
the pattern required.

If the scenario would have other like replacing \verb!r₅! by
\begin{verbatim}
  cnf(r₅, r ∨ p ∨ q, ...
\end{verbatim}
The \resolve rule have could derive $(p ∨ q) ∨ (p ∨ q)$, but again, that is not the expected result.
\end{myexamplenum}

Therefore, we perform a sequence of rearrangements inside the
involved formulas to match with the expected pattern by the
\emph{resolve} inference rule in Fig.~\ref{fig:metis-inferences}.

@@

\Needspace{3\baselineskip}
\begin{proof}
  \begin{equation*}
  \begin{bprooftree}
    \AxiomC{$Γ ⊢ φ₁$}
    \RightLabel{Lemma~\ref{lem:reorder-and-or}}
    \UnaryInfC{$Γ ⊢ \freorder_{∨}~φ₁~(ℓ ∨ ψ)$}
    \RightLabel{Lemma~\ref{lem:reorder-and-or}}
    \AxiomC{$Γ ⊢ φ₂$}
    \UnaryInfC{$Γ ⊢ \freorder_{∨}~φ₂~(¬ ℓ ∨ ψ)$}
    \RightLabel{∧-intro}
    \BinaryInfC{$Γ ⊢ \freorder_{∨}~φ₁~(ℓ ∨ ψ) ∧ \freorder_{∨}~φ₂~(¬ ℓ ∨ ψ)$}
    \RightLabel{Lemma~\ref{lem:rsol}}
    \UnaryInfC{$Γ ⊢ \frsol~(\freorder_{∨}~φ₁~(ℓ ∨ ψ) ∧
     \freorder_{∨}~φ₂~(¬ ℓ ∨ ψ))$}
    \RightLabel{by (\ref{eq:resolve})}
    \UnaryInfC{$Γ ⊢ \fresolve~φ₁~φ₂~ℓ~ψ$}
  \end{bprooftree}
  \end{equation*}
\end{proof}

>>>>

% -- TODO algun texto aqui?

@@

A \emph{literal} is a propositional variable (positive literal) or a
negation of a propositional variable (negative literal).

@@

\begin{notation}
\Lit is a synonym of \Prop to refer to literals.
\end{notation}

@@

The \emph{conjunctive normal form} of a formula also called clausal
normal form is the logical equivalent version expressed as a
conjunction of clauses where a \emph{clause} is the disjunction of
zero or more literals.
\end{mydefinition}

% ----------------------------------------------------------------------------
% -- RESOLVE --
% ----------------------------------------------------------------------------

The \resolve inference rule in combination with the \simplify rule is  the
resolution theorem for propositional logic that we found in the \Metis \TSTP
derivations (see this rule in~Fig.~\ref{fig:metis-inferences}). In particular,
to reconstruct the \resolve rule following the pattern in
Example~\ref{ex:inference-rule-pattern}, it is necessary to have a mechanism to
test for logic equivalence between propositions. The challenge is because of we
left out the semantics to treat only the syntax aspects of the propositional
formulas, therefore we have to manipulate in somehow the formulas to see if they
are logically equivalent.

Our approach to address this issue mainly consists of checking for syntactical
equality between the formulas. Firstly by converting them to their conjunctive
normal and secondly, by performing reordering of the inner formulas of the
resulting normal forms in the first conversion when they were not equal.

In~\cite{Prieto-Cubides2017a}, we make a description of the complete machinery
we use for that purpose. For instance, we define functions of the type ($\Source
\to  \Target \to \Prop$) like $\freorder_{∨}$ to fix the order of a disjunction
(premise) given by the order in the literals of another disjunction (conclusion)
if it is possible. Following the same way, we define for conjunctions the
function $\freorder_{\wedge}$ and for conjunctive normal forms the function
$\freorder_{\wedge\vee}$.

@@
